---
title: "bird_calcium"
author: "kevil khadka"
subtitle: stat_267
output:
  pdf_document: default
  html_document: default
---

This R markdown is the placeholder for your R codes you would use during the exam. You should write your final answers on the provided exam sheets.

R codes in R markdown are organized in multiple chunks starting with three-ticks and the letter r and end with another three-ticks. The following R chunk is for loading the `tidyverse` packages

```{r}
library(tidyverse)
```

You can execute this R chunk line-by-line by move the cursor to the line of desired and hit Ctrl-Enter, or execute the whole chunk by clicking on the play button (triangular shape) on the upper-right corner of the chunk.

For the problem 8, the data set is inputed as:

```{r}
bird <- tibble(
  calcium = c(120, 124, 111, 115, 110, 
              117, 104, 103, 116, 100, 
              150, 142, 133, 155, 160, 
              150, 138, 145, 140, 147), 
  condition = rep(c("female control", 
                    "male control", 
                    "female treatment", 
                    "male treatment"), each = 5))
View(bird)
```

part (a)

```{r}
bird %>% 
  group_by(condition) %>% 
  summarise(group_mean = mean(calcium), 
            group_sd = sd(calcium))

ggplot() + geom_boxplot(data=bird, aes(x = condition, y = calcium))
```
# doesn't seem like there is any outlier.But, it looks like the SDs are not equal across different groups. So, we may need to transform the data. 

part (b)

```{r}
mod <- lm(calcium ~ condition, data = bird)
anova(mod)
aov(mod)
```
## Here, P-value < 0.05. So, we reject our null hypothesis. It conclude that for birds, their body's ability to use calcium depends on the level of certain sex-related hormones in the blood.

```{r}
## Use a “quick and dirty” method
n_condtion <- count(bird, condition_number = bird$condition)
n_condtion
```
 
## Use a “quick and dirty” method
# formula SE = (sq_root of MSE) / (sq_root of # of observation in each group)

SE_f_con = (59.25)^1/2 / (5)^1/2  = 3.44
SE_f_tm = (59.25)^1/2 / (5)^1/2 = 3.44
SE_m_con = (59.25)^1/2 / (5)^1/2 = 3.44
SE_m_tm = (59.25)^1/2 / (5)^1/2  = 3.44

SE for difference = (MSE)^1/2 * (2/n)^1/2 = (59.25)^1/2 * (2/5)^1/2 = 5.925
2 * S.E = 11.85

(when Mean < 2*SE , they are equal)
so, pair of male control - female control is equal 
and, pair of male-treatment and female-treatment is equal. 

## Fisher's LSD

MSE = 59.25
SD = 7.69
#for balance design, 
SE for difference = (MSE)^1/2 * (2/n)^1/2 = (59.25)^1/2 * (2/5)^1/2 = 12.1706
t-value = qt(0.975, 16) = 2.119905
yardstick = (t-value) * SE_for_difference = 2.119 * 12.1706 = 25.789

pair of male control - female control is equal, and
pair of male-treatment and female-treatment is equal. 

## Use the Tukey’s HSD

yardstick = q * SE_for_difference = 2.861 * 4.86 = 13.904
q <- qtukey(0.95, 4, 16)/sqrt(2) = 2.861


```{r}
mod2 <- aov(mod)
TukeyHSD(mod2)
```

so here, 
pair of male control - female control is equal and,
pair of male-treatment and female-treatment is equal. 